---
title: "Initial Results for Table 3 and 5"
output:
  html_document:
    df_print: paged
Author: Anagha Uppal
Date: Feb 2021
Purpose: Tables 3 & 5 for SES 2nd Paper
Inputs: Participant Data and SES Data
Outputs: Results
To do: find out why inc.desc is so bad (it's because it's a multinomial parading as a binomial), try replacing inc with edu in some places, Automate greenness vars, Continue
  table 4 and 5
---

#### These are sample results for Table 3 - personal (parcel NDVI) and local (participant 100m ndvi) level greenness. Table 3 personal level is defined by dependence of greenness metric, while local level is defined by income or education (SES) as dependent. Stratification samples for table 5 are by gender and race. In progress is a automated function for discovery to try many greenness metrics one-by-one.
#### One reason for the underwhelming results for local level could be that income model uses family binomial, whereas this is a multinomial categorical variable. 
#### Another step is to include visualizations of observed vs fitted.

Import libraries and grab data

```{r echo=FALSE,message=FALSE}
library(leaps)
library(nnet) #this contains func multinom through which we can model a categorical (non-binary) y (dependent) variable
library(tidyverse)
library(MASS) #for stepAIC
library(GGally)
library(bootStepAIC) #for bootstrapping
library(plotly) #for bootstrap viz
library(pscl) #for pR2
library(styler) #style Rmd
library(corrP) #associations between all independent vars
library(splines) #for additive model
library(jtools) #for viz predictions
library(earth) #for masss lm
library(dplyr)

# styler::style_file()

# Grab the data
participants <- read.csv("C:/Users/a0uppa01/Documents/GIS/GHData_AU_20201221.csv")
p1 <- read.csv("C:/Users/a0uppa01/Documents/GIS/GHData_RY_041720.csv")
greenness <- read.csv("C:/Users/a0uppa01/Documents/GIS/participants_allgreenness_20210309.csv")
addparcelinfo <- read.csv("C:/Users/a0uppa01/Documents/GIS/GHParticipants_newrentown.csv")
```

Data Preparation

```{r echo=FALSE}

participants <- participants %>%
  mutate(gend.desc = as.factor(if_else(gend == 1, 'Female', 'Male'))) %>%
  mutate(race.desc = as.factor(if_else(race == 3, 'Black', 
                                       if_else(race == 5, 'White', 'Other')))) %>% 
  mutate(ethn.desc = as.factor(if_else(ethn == 1, 'Hispanic/Latino', 'Non Hispanic/Latino'))) %>%
  #mutate(shemp.desc = as.factor(if_else(shemp == 1, 'Yes', 'No'))) %>%
  mutate(inc.desc = as.factor(if_else(inc == 1, '< $20,000', 
                                      if_else(inc == 2, '20,000-$45,000', 
                                              if_else(inc == 3, '$45,000-$65,000',
                                                      if_else(inc==4 | inc==5 | inc==6, '>$65,000', NULL)))))) %>%
  mutate(edu.desc = as.factor(if_else(edu>=1 & edu<=3, '<=High School Diploma',
                                      if_else(edu==4 | edu==5, 'Some college',
                                              if_else(edu>=6 & edu<=8, '>=4-year degree', NULL))))) %>%
  mutate(medu.desc = as.factor(if_else(medu>=1 & medu<=3, '<=High School Diploma',
                                       if_else(medu==4 | medu==5, 'Some college',
                                               if_else(medu>=6 & medu<=8, '>=4-year degree', NULL))))) %>%
    mutate(rentown.desc = as.factor(if_else(rentown == 1, 'Rent', 'Own'))) %>%
  mutate(NEAR_DIST_MajorRoads = NEAR_DIST_MajorRoads/3281)

addparcelinfo[addparcelinfo==""]<-NA


participants <- left_join(participants, addparcelinfo, by=c("stid"="StudyID"))
participants <- participants %>% dplyr::select(inc.desc, stid, age, gend.desc,
                                               race.desc, ethn.desc, edu.desc,
                                               NEAR_DIST_MajorRoads,PopDen,
                                               CUR_TOTAL, PROP_CLASS, rentown_parcel,hznum,medu.desc)
# participants <- participants %>% dplyr::select(inc.desc, stid, age, gend.desc,
#                                                race.desc, ethn.desc, edu.desc,
#                                                NEAR_DIST_MajorRoads,rentown.desc)
# participants <- participants %>% dplyr::select(inc.desc, stid, age, gend.desc,
#                                                race.desc, ethn.desc, edu.desc,
#                                                NEAR_DIST_MajorRoads)

participant_greenness  <- left_join(participants, greenness, by=c("stid"="StudyID"))


# taking all variables from SES table 
# checking the NA count for each
# and removing rentown.desc and bc.desc (type of building) 
# these two variables have n=300 missing values
drop <- c("Nobuilding_NDVI_L19_Sum","lessvisible_NDVI_L19_MEAN", 
          "morevisible_NDVI_L19_MEAN", "Nobuilding_NDVI_L19_Mean", "Parcel_NDVI_L19_Sum",
          "Parcel_NDVI_L19_Mean","lessvisible_LAI_MEAN", "morevisible_LAI_MEAN",
          "morevisible_leafarea_MEAN", "lessvisible_leafarea_MEAN", "lessvisible_ndvi_MEAN",
          "morevisible_biomass_MEAN", "lessvisible_biomass_MEAN", 
          "SubjectID",
          "Field", "Event_Name", "Study_ID", "stid",
          "USER_Home_Address", "OBJECTID_1", "X", "Y", "ObjectID",
          "Subject_ID", "Home_Address", 
          "ImperviousyPercFocal500m060320", "ImperviousyPercFocal300m060320",
          "ImperviousyPercFocal50m060320",
          "MEAN_12", "Event_Name", "OBJECTID", "lessvisible_canopy_MEAN")
pg_for_lm = participant_greenness[,!(names(participant_greenness) %in% drop)]
pg_for_lm = pg_for_lm %>% dplyr::select(-starts_with("HEAL"),
                                 -starts_with("Heal"),
                                 -starts_with("FocalStatistics"),
                                 -starts_with("Andrea"),
                                 -starts_with("ghKDE"),
                                 -starts_with("ParcelID"),
                                 -starts_with("FIPS"),
                                 -starts_with("NEAREST")) 
# sapply(pg_for_lm, function(x) sum(is.na(x)))


                                      
pg_for_lm <- na.omit(pg_for_lm) #641 observations

pg_for_lm_owners <- filter(pg_for_lm, rentown_parcel == "Own")

```

```{r}
# pg_for_lm <- read.csv("C:/Users/a0uppa01/Documents/GIS/pg_for_lm_AU_20201221.csv", stringsAsFactors = T)
# pg_for_lm$inc.desc <- as.factor(pg_for_lm)
```


#### SUBSET Data Preparation  
A personal and local set of greenness data  
Greenness data consists of NDVI, Landsat NDVI, Leaf Area, Leaf Area Index, Biomass, and Tree Canopy at the following scales: Participant home radius of 20m, 50m, 100m, 150m, 200m, 300m, and 500m; Parcel-area, Parcel without the building, Backyard (named Less Visible), Frontyard (named More Visible); and Block Group  
Personal-level greenness refers to 20m around the Participant home and all parcel characteristics.
Area-level greenness refers to 50-500m greenness and Block group greenness.
Personal and area greenness will be assessed separately because participants may have more control over personal greenness than area greenness.   

```{r}
pg_for_lm_local = pg_for_lm %>% select(-starts_with("Participant_20m"), 
                                           -starts_with("Parcel_"),
                                           -starts_with("Nobuilding_"),
                                           -starts_with("lessvisible_"),
                                           -starts_with("morevisible_"))

pg_for_lm_personal = pg_for_lm %>% select(-contains("00m_"), 
                                              -contains("50m_"),
                                              -starts_with("Blockgroup_"))

```

##### Trying stepwise regression on full model, followed by bootstrapping:

```{r echo=FALSE}


```

##### Bootstrapping viz (skipping; problems in code)

```{r echo=FALSE, eval=FALSE}
# Base plot for number of times a variable was picked by stepAIC
plot_ly(coef.pick, x = rownames(coef.pick), y = pick,
        type = "bar", opacity = 0.75, name = "Times picked (%)",
        hoverinfo = "text", text = pick.text,
        marker = list(color = "#00994d", line = list(width = 2))) %>% 
  
  # Layer for number of times a variable was statistically significant at 5%
  add_trace(data = coef.stat.long, x = variable, y = sig, 
            type = "scatter", mode = "markers + line", name = "Stat. Sig (%)",
            line = list(color = "#ffdb4d", width = 15),
            hoverinfo = "text", text = text) %>% 
  
  # Layer for number of times a variable's coefficient was positive
  add_trace(data = coef.sign, x = rownames(coef.pick), y = rep(-5, nCovariates), 
            type = "scatter", mode = "markers", name = "Coef Sign(% pos)",
            marker = list(symbol = "triangle-up", size = pos/scale, color = "#4da6ff",
                          line = list(color = "black", width = 2)),
            hoverinfo = "text", text = sign.text.up) %>% 
  
  # Layer for number of times a variable's coefficient was negative
  add_trace(data = coef.sign, x = rownames(coef.pick), y = rep(-10, nCovariates), 
            type = "scatter", mode = "markers", name = "Coef Sign(% neg)",
            marker = list(symbol = "triangle-down", size = neg/scale, color = "#ff704d",
                          line = list(color = "black", width = 2)),
            hoverinfo = "text", text = sign.text.down) %>% 
  
  # Layout, annotations, axis options etc
  layout(xaxis = list(title = "<b>Covariates</b>"),
         yaxis = list(title = "<b>Percentage(%)</b>",
                      tickmode = "array", 
                      tickvals = round(seq(0, 100, length.out = 10), 0),
                      domain = c(0.2, 1)),
         plot_bgcolor = "#e1efc3",
         paper_bgcolor = "#e1efc3",
         
         annotations = list(
           list(x = 0.1, y = 1, 
                xref = "paper", yref = "paper", 
                xanchor = "left", yanchor = "top",
                ax = 0, ay = 0,
                text = "Visualizing <em>boot.stepAIC()</em>",
                font = list(family = "serif", size = 30)),
           
           list(x = 0.3, y = 0.1, 
                xref = "paper", yref = "paper", 
                xanchor = "left", yanchor = "top",
                ax = 0, ay = 0,
                text = paste("<em>Original Model:</em>", origModel),
                font = list(family = "PT Sans Narrow", size = 15)),
           
           list(x = 0.21, y = 0.05, 
                xref = "paper", yref = "paper", 
                xanchor = "left", yanchor = "top", align = "left",
                ax = 0, ay = 0,
                text = paste("<em>Stepwise Model:</em>", stepModel),
                font = list(family = "PT Sans Narrow", size = 15)),
           
           list(x = 0.8, y = 0.90, 
                xref = "paper", yref = "paper", 
                xanchor = "left", yanchor = "top", align = "left",
                ax = 0, ay = 0,
                text = paste0("<em>No. of Covariates:</em>", nCovariates, "<br>",
                              "<em>No. of bootstrap samples:</em>", nBoot, "<br>"),
                font = list(family = "PT Sans Narrow", size = 15))
         ))
```

  
    

Breaking up the analysis now, we separate the analysis of personal-level 
greenness with local-level greenness.  
Personal greenness is defined as greenness at 20m radius or parcel greenness.   
We run linear models both to estimate personal greenness from SES variables as well as to estimate income or education from personal greenness and other SES variables.  

First, let's check correlations between some SES variables
```{r echo=FALSE}
pg_for_corr <- pg_for_lm %>% select(age, race.desc, ethn.desc, gend.desc,
                                    inc.desc, edu.desc)
pg_corr = corrP(pg_for_corr,parallel = TRUE, n.cores = 4, p.value = 0.05)
corrplot::corrplot(pg_corr)
corrgram::corrgram(pg_corr)
```
0.4 corr between race and ethnicity -- nothing surprising and nothing to greatly worry about.

##### Personal-level assessment -   
Greenness <- SES  
Greenness is dependent on SES variables  
For now, we have chosen parcel to represent greenness  

```{r echo=FALSE}
# PG = Personal Greenness with Interactions

pgi.model <- glm(Parcel_ndvi_Mean ~(inc.desc+edu.desc+age+gend.desc+race.desc)^2, 
                  data = pg_for_lm)
summary(pgi.model)
```
We see that interactions between low income, education and race(white) are significant
  
  \
```{r}

```
  

##### Personal greenness without interactions
```{r echo=FALSE}
pg.model <- glm(Parcel_ndvi_Mean ~inc.desc+edu.desc+age+gend.desc+race.desc, 
                  data = pg_for_lm)
summary(pg.model)
```
Low income continues to be highly (negatively) significant.  
  
##### Personal greenness stratified by gender
```{r echo=FALSE}
pgim.model <- glm(Parcel_ndvi_Mean ~(inc.desc+edu.desc+age+race.desc)^2, 
                  data = pg_for_lm, 
                  subset= gend.desc=="Male")
summary(pgim.model)
pgif.model <- glm(Parcel_ndvi_Mean ~(inc.desc+edu.desc+age+race.desc)^2, 
                      data = pg_for_lm, 
                      subset= gend.desc=="Female")
summary(pgif.model)

pgm.model <- glm(Parcel_ndvi_Mean ~inc.desc+edu.desc+age+race.desc, 
                  data = pg_for_lm, 
                  subset= gend.desc=="Male")
summary(pgm.model)
pgf.model <- glm(Parcel_ndvi_Mean ~inc.desc+edu.desc+age+race.desc, 
                      data = pg_for_lm, 
                      subset= gend.desc=="Female")
summary(pgf.model)
```
Income, education and race are all significant interaction variables, with income being the most important.      
  
##### Let's try including only a few interaction variables
```{r echo=FALSE}
pgii.model <- glm(Parcel_ndvi_Mean ~inc.desc+edu.desc+age+gend.desc+race.desc+
                    inc.desc*edu.desc+inc.desc*race.desc+edu.desc*race.desc, 
                  data = pg_for_lm)
summary(pgii.model)
```
  
  
  
##### Personal-level assessment -   
SES <- Greenness+SES  
Income or education is dependent on greenness and other SES variables

```{r echo=FALSE, message=FALSE}
# Income AND greenness - multivariate 
pgmult.model <- glm(cbind(inc.desc,edu.desc) ~Parcel_ndvi_Mean+age+
                               race.desc+gend.desc+
                               NEAR_DIST_MajorRoads, 
                  data = pg_for_lm, family=binomial())
# Stepwise regression model
step.model <- stepAIC(pgmult.model, direction = "both", trace = FALSE)
bootstep.model <- boot.stepAIC(pgmult.model, pg_for_lm, B = 100, alpha = 0.05, 
                               direction = "both", k = 2, verbose = FALSE)
summary(pgmult.model)
summary(step.model) 
summary(bootstep.model)
bootstep.model

anova(pgmult.model, test="Chisq")
# McFadden's pseudo-R2
# pR2(pgmult.model)
car::vif(pgmult.model) 

```
NO variables found to be significant. Best model null model ..?  


```{r echo=FALSE, eval=FALSE}
pgmult1.model <- glm(cbind(inc,edu) ~Parcel_ndvi_Mean+age+
                               race+gend+
                               NEAR_DIST_MajorRoads, 
                  data = pg_for_lm, family=binomial())
```
Throws error  

##### Just one SES dependent var

```{r echo=FALSE}
pgsesi.model <- glm(inc.desc ~(Parcel_ndvi_Mean+age+
                               race.desc+ethn.desc+gend.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial())
summary(pgsesi.model)

pgses.model <- glm(inc.desc ~Parcel_ndvi_Mean+age+
                               race.desc+ethn.desc+gend.desc+
                               NEAR_DIST_MajorRoads, 
                  data = pg_for_lm, family=binomial())
summary(pgses.model)

```
Nothing  
  
  
  ##### Stratified samples

```{r echo=FALSE}
pgsesim.model <- glm(inc.desc ~(Parcel_ndvi_Mean+age+
                               race.desc+ethn.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial(),
                  subset= gend.desc=="Male")
summary(pgsesim.model)

pgsesiw.model <- glm(inc.desc ~(Parcel_ndvi_Mean+age+
                               ethn.desc+gend.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial(),
                  subset= race.desc=="White")
summary(pgsesiw.model)

```
Age+Near_dist  
  
##### Checks for personal greenness model
```{r echo=FALSE}
car::vif(pgi.model) 
alias(pgi.model) #Based on this, had to remove ethnicity
# Exclude interactions or do something else?
anova(pgi.model, test="Chisq")
alias(pgi.model) #Based on this, had to remove ethnicity
# McFadden's pseudo-R2
pR2(pgi.model)
```
  
  
##### Local-level assessment -   
SES <- Greenness+SES  
Income or education is dependent on greenness and other SES variables  
Greeness var changes to local level  

```{r}

lgi.model <- glm(inc.desc ~(Participant_100m_Mean_ndvi+age+
                               race.desc+gend.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial())
summary(lgi.model)

lg.model <- glm(inc.desc ~Participant_100m_Mean_ndvi+age+
                               race.desc+gend.desc+
                               NEAR_DIST_MajorRoads, 
                  data = pg_for_lm, family=binomial())
summary(lg.model)

lgim.model <- glm(inc.desc ~(Participant_100m_Mean_ndvi+age+
                               race.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial(), 
                  subset= gend.desc=="Male")
summary(lgim.model)
lgif.model <- glm(inc.desc ~(Participant_100m_Mean_ndvi+age+
                               race.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial(), 
                  subset= gend.desc=="Female")
summary(lgif.model)

lgiw.model <- glm(inc.desc ~(Participant_100m_Mean_ndvi+age+
                               ethn.desc+gend.desc+
                               NEAR_DIST_MajorRoads)^2, 
                  data = pg_for_lm, family=binomial(), 
                  subset= race.desc=="White")
summary(lgiw.model)

car::vif(lgim.model) 
alias(lgim.model) #Based on this, had to remove ethnicity
# Exclude interactions or do something else?
anova(lgim.model, test="Chisq")
alias(lgim.model) #Based on this, had to remove ethnicity
# McFadden's pseudo-R2
# pR2(lgim.model)



```
Age*Near_dist found for all stratified samples  


```{r eval=FALSE}

glmnet()

```
  

##### Bootstrapping Visualization (some errors need to be resolved)
```{r echo=FALSE, eval=FALSE}
###############################################################
## Visualization
fit.boot <- bootstep.model
nBoot <- summary(fit.boot)[8,1]
origModel <- paste(names(coef(fit.boot$OrigModel)), collapse = " + ")
stepModel <- paste(names(coef(fit.boot$OrigStepAIC)), collapse = " + ")
# Names of covariates
covariates <- rownames(fit.boot$Covariates)
nCovariates <- length(covariates)
# Matrix of number of times each covariate was picked
coef.pick <- fit.boot$Covariates
# Matrix for the consistency of sign on each covariate
coef.sign <- fit.boot$Sign
# Change names since many are factors
rownames(coef.sign)[1] <- "exr.desc"
rownames(coef.sign)[5] <- "inc.desc"
rownames(coef.sign)[6] <- "saf3.desc"
rownames(coef.sign)[7] <- "edu.desc"
rownames(coef.sign)[22] <- "inc.desc"
rownames(coef.sign)[23] <- "inc.desc"


coef.sign <- coef.sign[match(rownames(coef.pick), rownames(coef.sign)),]
# Matrix for statistical significance
coef.stat <- fit.boot$Significance
# Change name for "chas" since it is a factor

rownames(coef.stat)[16] <- "inc.desc"
rownames(coef.stat)[17] <- "obese.desc"
rownames(coef.stat)[21] <- "race.desc"
rownames(coef.stat)[22] <- "inc.desc"
rownames(coef.stat)[23] <- "inc.desc"
rownames(coef.stat)[24] <- "saf1.desc"
rownames(coef.stat)[25] <- "medu.desc"
rownames(coef.stat)[26] <- "edu.desc"


coef.stat <- coef.stat[match(rownames(coef.pick), rownames(coef.stat)),]
# Make into long form for charting later
coef.stat.long <- data.frame()
for(i in 1:length(coef.stat)){
  n <- round(coef.stat[i],0)
  vec <- seq(0, n, by = 2)
  mat <- data.frame(rep(names(coef.stat)[i], length(vec)), vec, paste("% Sig", n))
  names(mat) <- c("variable", "sig", "text")
  
  # We'll use mode = "line". NA helps separate line segments
  coef.stat.long <- rbind(coef.stat.long, mat, c(NA, NA))
}
# Convert to dataframes
coef.pick <- as.data.frame(coef.pick)
coef.stat <- as.data.frame(coef.stat)
coef.sign <- as.data.frame(coef.sign)
names(coef.pick) <- "pick"
names(coef.sign) <- c("pos", "neg")
names(coef.stat) <- "stat"

```

  
#####  PG = Personal Greenness with Additive
```{r}
# e

pga.model <- glm(Parcel_ndvi_Mean ~(inc.desc+edu.desc+bs(age)+gend.desc+race.desc)^2, 
                  data = pg_for_lm)
summary(pga.model)
#effect_plot(pga.model, pred=inc.desc, interval = TRUE, plot.points = TRUE)

car::vif(pga.model) 
alias(pga.model) #Based on this, had to remove ethnicity
# Exclude interactions or do something else?
anova(pga.model, test="Chisq")
alias(pga.model) #Based on this, had to remove ethnicity
# McFadden's pseudo-R2
pR2(pga.model)
```

  
    
```{r}
reg=earth(Parcel_ndvi_Mean ~(inc.desc+edu.desc+age+gend.desc+race.desc)^2,data=pg_for_lm)
summary(reg)
```
##### Local greenness with interaction
```{r}
pgi.model <- glm(Participant_100m_Mean_ndvi ~(inc.desc+edu.desc+age+gend.desc+race.desc)^2, 
                  data = pg_for_lm)
summary(pgi.model)
```


##### Local greenness without interactions
```{r echo=FALSE}
pg.model <- glm(Participant_100m_Mean_ndvi ~inc.desc+edu.desc+age+gend.desc+race.desc, 
                  data = pg_for_lm)
summary(pg.model)
```
Low income continues to be highly (negatively) significant.  
  
##### Local greenness stratified by gender
```{r echo=FALSE}
pgim.model <- glm(Participant_100m_Mean_ndvi ~(inc.desc+edu.desc+age+race.desc)^2, 
                  data = pg_for_lm, 
                  subset= gend.desc=="Male")
summary(pgim.model)
pgif.model <- glm(Participant_100m_Mean_ndvi ~(inc.desc+edu.desc+age+race.desc)^2, 
                      data = pg_for_lm, 
                      subset= gend.desc=="Female")
summary(pgif.model)

pgm.model <- glm(Participant_100m_Mean_ndvi ~inc.desc+edu.desc+age+race.desc, 
                  data = pg_for_lm, 
                  subset= gend.desc=="Male")
summary(pgm.model)
pgf.model <- glm(Participant_100m_Mean_ndvi ~inc.desc+edu.desc+age+race.desc, 
                      data = pg_for_lm, 
                      subset= gend.desc=="Female")
summary(pgf.model)
```
Income, education and race are all significant interaction variables, with income being the most important.      

